Point-process based representation learning for Electronic Health Records

### 2023-03-06 Manuscript_v1.tex
update

For each desired patient, we can compute the mean of cosine similarity between the 10 nearest neighbors (in the embedding space) and the desired patients.
\subsection*{Training Details}
To be completed.

@@

% Table generated by Excel2LaTeX from sheet 'time concat'
\begin{table*}[htbp]
    \centering
    \caption{Add caption}
      \begin{tabular}{cccccccccccccc}
      \toprule
      \toprule
            &       & \multicolumn{8}{c}{TEEDAM}                                    &       &       &       &  \\
  \cmidrule{3-10}          &       & \multicolumn{2}{c}{AE (next mark)} &       & \multicolumn{2}{c}{PP(single+mark)} &       & \multicolumn{2}{c}{PP (MC/ML)} &       &       &       &  \\
  \cmidrule{3-4}\cmidrule{6-7}\cmidrule{9-10}    Dataset & Metric & concat & sum   &       & concat & sum   &       & concat & sum   & Latent & SAHP  & THP   & GRU-CP \\
      \midrule
      \multirow{2}[2]{*}{SO (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -0.56 & -0.57 &       & -2.04 & -2.05 & -1.54 & -1.86 & -1.84 & \textit{NR} \\
            & F1-score & \textbf{38.46} & 36.91 &       & 36.04 & 34.91 &       & 32.65 & 31.67 & 28.34 & 24.12 & 23.89 & 26 \\
      \midrule
      \multirow{2}[2]{*}{ReTweet (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -6.91 & -8.036 &       & -11   &       & -3.89 & -4.56 & -4.57 & \textit{NR} \\
            & F1-score & \textbf{62.69} & 61.39 &       & 53.91 & 38.63 &       & 53    &       & 58.29 & 53.92 & 53.86 & \textit{NR} \\
      \midrule
      \multirow{2}[2]{*}{Synthea (ML)} & LL/\#events & \textit{ND} & \textit{ND} &       & -2.405 & -2.599 &       & -7.254 & -6.58 & \textit{ND} & \textit{ND} & \textit{ND} & \textit{NR} \\
            & AUROC & \textbf{89.58} & 89.19 &       & 64.98 & 63.95 &       & 60.95 & 60.65 & \textit{ND} & \textit{ND} & \textit{ND} & 0.85(.014) \\
      \midrule
      \multirow{2}[2]{*}{ReTweet (ML)} & LL/\#events & \textit{ND} & \textit{ND} &       & 2.045 & 1.863 &       & -1.517 & -1.601 & \textit{ND} & \textit{ND} & \textit{ND} & \textit{NR} \\
            & AUROC & 69.63 & 68.54 &       & 69.31 & 68.05 &       & \textbf{74.22} & 71.27 & \textit{ND} & \textit{ND} & \textit{ND} & 0.611(0.001) \\
      \bottomrule
      \bottomrule
      \end{tabular}%
    \label{tab:1}%
  \end{table*}%

@@

% Table generated by Excel2LaTeX from sheet 'Sheet3'
\begin{table*}[htbp]
    \centering
    \caption{Add caption}
      \begin{tabular}{ccccccccccc}
      \toprule
      \toprule
            &       &       & \multicolumn{2}{c}{F1} &       & \multicolumn{2}{c}{AUPRC} &       & \multicolumn{2}{c}{AUROC} \\
  \cmidrule{4-5}\cmidrule{7-8}\cmidrule{10-11}    Dataset & Setting & Center & DAM   & TE+DAM &       & DAM   & TE+DAM &       & DAM   & TE+DAM \\
      \midrule
      \multirow{8}[6]{*}{P12} & \multirow{3}[2]{*}{mc1} & 1     & 0.53 (0.04) & 0.52 (0.03) &       & 0.53 (0.06) & 0.53 (0.04) &       & 0.85 (0.03) & 0.85 (0.03) \\
            &       & 2     & 0.56 (0.02) & 0.55 (0.03) &       & 0.56 (0.05) & 0.56 (0.03) &       & 0.86 (0.01) & 0.87 (0.01) \\
            &       & 3     & 0.55 (0.04) & \textbf{0.57 (0.03)} &       & 0.59 (0.05) & 0.58 (0.03) &       & 0.86 (0.02) & 0.87 (0.01) \\
  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc2} & 1     & 0.52  & 0.52  &       & 0.53  & 0.52  &       & 0.85  & 0.85 \\
            &       & 2     & 0.54  & 0.54  &       & 0.53  & \textbf{0.55} &       & 0.85  & 0.86 \\
            &       & 3     & 0.52  & 0.54  &       & 0.56  & 0.56  &       & 0.85  & 0.86 \\
  \cmidrule{2-11}          & sc    & -     & \textbf{0.49 (0.05)} & 0.46 (0.01) &       & \textbf{0.48 (0.05)} & 0.44 (0.02) &       & 0.82 (0.03) & 0.81 (0.02) \\
            & seft  & -     & \textbf{0.52 (0.03)} & 0.5 (0.03) &       & \textbf{0.51 (0.05)} & 0.47 (0.03) &       & 0.83 (0.01) & 0.82 (0.01) \\
      \midrule
      \multirow{8}[6]{*}{P19} & \multirow{3}[2]{*}{mc1} & 1     & 0.53 (0.03) & 0.52 (0.04) &       & 0.51 (0.05) & \textbf{0.53 (0.05)} &       & 0.84 (0.03) & 0.84 (0.02) \\
            &       & 2     & 0.51  & 0.51  &       & 0.51  & 0.51  &       & 0.85  & 0.85 \\
            &       & 3     & 0.7 (0.02) & 0.69 (0.01) &       & 0.78 (0.02) & 0.78 (0.02) &       & 0.92 (0.02) & 0.93 (0.01) \\
  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc2} & 1     & 0.62 (0.04) & 0.62 (0.03) &       & 0.7 (0.03) & 0.69 (0.03) &       & 0.94 (0.01) & 0.94 (0.01) \\
            &       & 2     & 0.45  & \textbf{0.59} &       & 0.68  & 0.67  &       & 0.88  & 0.87 \\
            &       & 3     & 0.62  & 0.61  &       & 0.53  & \textbf{0.55} &       & 0.92  & 0.91 \\
  \cmidrule{2-11}          & sc    & -     & 0.68 (0.02) & 0.68 (0.03) &       & 0.77 (0.02) & 0.76 (0.03) &       & 0.93 (0.0) & 0.93 (0.0) \\
            & seft  & -     & 0.61 (0.02) & 0.61 (0.01) &       & 0.68 (0.01) & 0.67 (0.01) &       & 0.94 (0.0) & 0.94 (0.0) \\
      \bottomrule
      \bottomrule
      \end{tabular}%
    \label{tab:3}%
  \end{table*}%

@@

Fig 1 visualizes the t-SNE plot for the learned embeddings of [] with and without event encodings. Besides, we visualize the sampling pattern of an example patient and its neighbors. As can be seen,  in TEEDAM neighbors have more similar pattern compared to the baseline. Although event encoding does not necessarily leads to better performance, we can learn a better representation that considers both absolute values and the patterns. One application could be time series generation.

>>>>

The mean of cosine similarities between the measurement density of a desired patient and its 10 nearest neighbors (in the embedding space) is computed (10-nnps).
\subsection*{Training Details}
To be completed.

@@

% Table generated by Excel2LaTeX from sheet 'time concat'
\begin{table*}[htbp]
    \centering
    \caption{Add caption}
      \begin{tabular}{cccccccccccccc}
      \toprule
      \toprule
            &       & \multicolumn{8}{c}{TEEDAM}                                    &       &       &       &  \\
  \cmidrule{3-10}          &       & \multicolumn{2}{c}{AE (next mark)} &       & \multicolumn{2}{c}{PP(single+mark)} &       & \multicolumn{2}{c}{PP (MC/ML)} &       &       &       &  \\
  \cmidrule{3-4}\cmidrule{6-7}\cmidrule{9-10}    Dataset & Metric & concat & sum   &       & concat & sum   &       & concat & sum   & Latent & SAHP  & THP   & GRU-CP \\
      \midrule
      \multirow{2}[2]{*}{SO (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -0.56 & -0.57 &       & -2.04 & -2.05 & -1.54 & -1.86 & -1.84 & \textit{NR} \\
            & F1-score & \textbf{38.46} & 36.91 &       & 36.04 & 34.91 &       & 32.65 & 31.67 & 28.34 & 24.12 & 23.89 & 26 \\
      \midrule
      \multirow{2}[2]{*}{ReTweet (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -6.91 & -8.036 &       & -11   & -23.38 & -3.89 & -4.56 & -4.57 & \textit{NR} \\
            & F1-score & \textbf{62.69} & 61.39 &       & 53.91 & 38.63 &       & 53    & 35.17 & 58.29 & 53.92 & 53.86 & \textit{NR} \\
      \midrule
      \multirow{2}[2]{*}{Synthea (ML)} & LL/\#events & \textit{ND} & \textit{ND} &       & -2.405 & -2.599 &       & -7.254 & -6.58 & \textit{ND} & \textit{ND} & \textit{ND} & \textit{NR} \\
            & AUROC & \textbf{89.58} & 89.19 &       & 64.98 & 63.95 &       & 60.95 & 60.65 & \textit{ND} & \textit{ND} & \textit{ND} & 0.85(.014) \\
      \midrule
      \multirow{2}[2]{*}{ReTweet (ML)} & LL/\#events & \textit{ND} & \textit{ND} &       & 2.045 & 1.863 &       & -1.517 & -1.601 & \textit{ND} & \textit{ND} & \textit{ND} & \textit{NR} \\
            & AUROC & 69.63 & 68.54 &       & 69.31 & 68.05 &       & \textbf{74.22} & 71.27 & \textit{ND} & \textit{ND} & \textit{ND} & 0.611(0.001) \\
      \bottomrule
      \bottomrule
      \end{tabular}%
    \label{tab:1}%
  \end{table*}%

@@

% Table generated by Excel2LaTeX from sheet 'Sheet3'
\begin{table*}[htbp]
    \centering
    \caption{Add caption}
      \begin{tabular}{ccccccccccc}
      \toprule
      \toprule
            &       &       & \multicolumn{2}{c}{F1} &       & \multicolumn{2}{c}{AUPRC} &       & \multicolumn{2}{c}{AUROC} \\
  \cmidrule{4-5}\cmidrule{7-8}\cmidrule{10-11}    Dataset & Setting & Center & DAM   & TE+DAM &       & DAM   & TE+DAM &       & DAM   & TE+DAM \\
      \midrule
      \multirow{10}[8]{*}{P12} & \multirow{3}[2]{*}{mc1} & 1     & 0.53 (0.04) & 0.52 (0.03) &       & 0.53 (0.06) & 0.53 (0.04) &       & 0.85 (0.03) & 0.85 (0.03) \\
            &       & 2     & 0.56 (0.02) & 0.55 (0.03) &       & 0.56 (0.05) & 0.56 (0.03) &       & 0.86 (0.01) & 0.87 (0.01) \\
            &       & 3     & 0.55 (0.04) & \textbf{0.57 (0.03)} &       & 0.59 (0.05) & 0.58 (0.03) &       & 0.86 (0.02) & 0.87 (0.01) \\
  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc2} & 1     & 0.52  & 0.52  &       & 0.53  & 0.52  &       & 0.85  & 0.85 \\
            &       & 2     & 0.54  & 0.54  &       & 0.53  & \textbf{0.55} &       & 0.85  & 0.86 \\
            &       & 3     & 0.52  & 0.54  &       & 0.56  & 0.56  &       & 0.85  & 0.86 \\
  \cmidrule{2-11}          & \multirow{3}[2]{*}{sc} & 1     & \textbf{0.49 (0.05)} & 0.46 (0.01) &       & \textbf{0.48 (0.05)} & 0.44 (0.02) &       & 0.82 (0.03) & 0.81 (0.02) \\
            &       & 2     & \textbf{0.52 (0.03)} & 0.5 (0.03) &       & \textbf{0.51 (0.05)} & 0.47 (0.03) &       & 0.83 (0.01) & 0.82 (0.01) \\
            &       & 3     & 0.53 (0.03) & 0.52 (0.04) &       & 0.51 (0.05) & \textbf{0.53 (0.05)} &       & 0.84 (0.03) & 0.84 (0.02) \\
  \cmidrule{2-11}          & seft  & -     & 0.51  & 0.51  &       & 0.51  & 0.51  &       & 0.85  & 0.85 \\
      \midrule
      \multirow{6}[6]{*}{P19} & \multirow{2}[2]{*}{mc1} & 1     & 0.7 (0.02) & 0.69 (0.01) &       & 0.78 (0.02) & \textbf{0.78 (0.02)} &       & 0.92 (0.02) & 0.93 (0.01) \\
            &       & 2     & 0.62 (0.04) & 0.62 (0.03) &       & 0.7 (0.03) & 0.69 (0.03) &       & 0.94 (0.01) & 0.94 (0.01) \\
  \cmidrule{2-11}          & \multirow{2}[2]{*}{mc2} & 1     & 0.45  & \textbf{0.59} &       & 0.68  & 0.67  &       & 0.88  & 0.87 \\
            &       & 2     & 0.62  & 0.61  &       & 0.53  & \textbf{0.55} &       & 0.92  & 0.91 \\
  \cmidrule{2-11}          & \multirow{2}[2]{*}{sc} & 1     & 0.68 (0.02) & 0.68 (0.03) &       & 0.77 (0.02) & 0.76 (0.03) &       & 0.93 (0.0) & 0.93 (0.0) \\
            &       & 2     & 0.61 (0.02) & 0.61 (0.01) &       & 0.68 (0.01) & 0.67 (0.01) &       & 0.94 (0.0) & 0.94 (0.0) \\
      \bottomrule
      \bottomrule
      \end{tabular}%
    \label{tab:3}%
  \end{table*}%

@@

Fig 1 visualizes the t-SNE plot for the learned embeddings of [] with and without event encodings. Besides, we visualize the sampling pattern of an example patient and its neighbors. As can be seen,  in TEEDAM neighbors have more similar pattern compared to the baseline. Although event encoding does not necessarily leads to better performance, we can learn a better representation that considers both absolute values and patterns.
The 10 nearest neighbor pattern similarity (10-nnps) of positive labels for TEE+DAM ($0.903 \pm 0.058 $) is signifcantly higher than (DAM) ($0.832 \pm 0.095$)($pvalue \leq 0.001$).
One application could be time series generation.
